\section{Discussion}

\textbf{Trade-offs.} ICL delivers strong alignment with small \(\Delta\) capability changes; PEFT maximizes alignment but incurs large negative \(\Delta\) (notably on Gemma-2 MMLU/GAIA); steering offers moderate alignment with trait-dependent \(\Delta\), improving with careful vector construction (e.g., purified openness). This suggests clear practical guidance: when preserving capability is critical, prefer ICL; use steering when lightweight, runtime control is needed and calibration is feasible; reserve PEFT for settings where strong, stable personality alignment outweighs capability costs. Detailed comparative analysis is in Appendix~\ref{app:comparative-analysis}.

\textbf{Personality Encoding Mechanisms.} Systematic manipulation reveals three key insights: (1) ICL's effectiveness with minimal disruption shows personality traits are accessible through surface-level conditioning without affecting core representations; (2) PEFT's strong alignment but capability costs demonstrate personality can be deeply encoded in parameters but at the expense of general capabilities, suggesting shared representational resources; (3) Activation steering's layer-specific effectiveness (typically Layer 15) maps personality traits to specific representational depths, showing personality encoding consolidates at intermediate transformer layers.

\textbf{Manipulation as Interpretability Tool.} Treating personality manipulation as controlled probes demonstrates that different interventions reveal complementary aspects of personality representation. The \(\Delta\)-analysis framework isolates method-specific effects and identifies which representational pathways each technique accesses, establishing manipulation-based interventions as a viable interpretability methodology for understanding behavioral control in neural language models.

\textbf{Trait/model dependence.} Agreeableness is hardest to align via ICL; openness benefits from vector composition; LLaMA-3 shows different absolute baselines across runs, so we interpret only within-run \(\Delta\). Complete trait-wise results are in Appendix~\ref{app:trait-results}.

\textbf{Limitations.} We avoid cross-run baseline comparisons by design (\(\Delta\)-only). We ignore \(S_{DIS}\) for BBQ and rely on the ambiguous bias score \(S_{AMB}\). Alignment is validated independently; detailed settings and per-subject analyses are in the appendices. Extended discussion of limitations, ethical considerations, and future work is in Appendix~\ref{app:discussion-extended}.